<!-- README.md is generated from README.Rmd. Please edit that file
knitr::knit('README.Rmd')
-->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
    )
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/mpio-be/rangeMapper/workflows/R-CMD-check/badge.svg)](https://github.com/mpio-be/rangeMapper/actions)
[![Coverage Status](https://img.shields.io/codecov/c/github/mpio-be/rangeMapper/master.svg)](https://codecov.io/github/mpio-be/rangeMapper?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/rangeMapper)](http://cran.r-project.org/package=rangeMapper)
[![cran checks](https://cranchecks.info/badges/worst/rangeMapper)](https://cran.r-project.org/web/checks/check_results_rangeMapper.html)
[![monthly](https://cranlogs.r-pkg.org/badges/rangeMapper)](https://www.rpackages.io/package/rangeMapper) 
[![total](https://cranlogs.r-pkg.org/badges/grand-total/rangeMapper)](https://www.rpackages.io/package/rangeMapper)
[![status](https://tinyverse.netlify.com/badge/mapview)](https://CRAN.R-project.org/package=mapview)

<!-- badges: end -->

[_rangeMapper_](http://onlinelibrary.wiley.com/doi/10.1111/j.1466-8238.2011.00739.x/full/) is a framework for the study of macroecological patterns of life-history traits.


### _rangeMapper_ in a nutshell

```{r,echo=FALSE,results='hide', message=FALSE,warning=FALSE, fig.width=10, fig.height=5}
    require(rangeMapper)
    require(magrittr)
    require(ggplot2)
    require(rgdal)

    breding_ranges = readOGR(system.file(package = "rangeMapper", "extdata", "wrens", "vector_combined"), 
                    "wrens", verbose = FALSE)
    breding_ranges = spTransform(breding_ranges, CRS('+proj=moll')) %>% rgeos::gBuffer(width = 0, byid = TRUE)
    wrens_boundary = rgeos::gUnionCascaded(breding_ranges)

    data(wrens)
    d = subset(wrens, select = c('sci_name', 'body_mass', 'clutch_size') )
    

    con = ramp("wrens.sqlite", gridSize = 150000, spdf = breding_ranges, biotab = d, ID = "sci_name",
                FUN = "median", overwrite = TRUE)


```
* Datasets (i.e. maps) are retrieved from `sqlite` project files as objects inheriting from `SpatialPixels` or `data.table`. Maps can be plotted directly with `plot()`.

```{r,1, fig.cap="", echo=TRUE,results='hide', message=FALSE,warning=FALSE, fig.width=10, fig.height=5}
    map = rangeMap.fetch(con, c('median_body_mass', 'median_clutch_size'), spatial = FALSE)
    plot(map, boundary = wrens_boundary, 
        ncol = 2 , 
        outlierDetector = function(x) extremevalues::getOutliersI(x)$limit, 
        colours = viridisLite::viridis(11) )



```

*  The link between the assemblage level (i.e. the raster cell) and the species level (i.e. the data behind the raster cell) is kept explicit at all stages of the project.
`MAP`s are constructed based on `SQL` aggregate functions or statistical models build in R and can be based on arbitrary subsets defined at both species and assemblage levels.

```R
rangeMap.save(con, FUN = lmSlope, biotab = "life_history_traits",
    biotrait = "body_mass", tableName = "slope_bm_cs",
    formula = log(body_mass) ~ clutch_size,
    subset = list(MAP_species_richness = "SR >= 5",
                  MAP_altitude = "altitude > 1000",
                  BIO_biotab = "Family = 'Troglodytidae'
                                 AND clutch_size is not NULL") )
```


### Installation
```R
remotes::install_github("mpio-be/rangeMapper") # development.
install.packages("rangeMapper") # stable.
```




